---
title: "Analysis"
author: "Babacar NDAO"
date: "12/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Supplemental analysis and cluster annotation

A- ALL GENES ANALYYSIS 


## -----------------------------------------------------------------------------
## Libraries
## -----------------------------------------------------------------------------

```{r}
suppressMessages(library(Seurat))
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(grid))
suppressMessages(library(S4Vectors))
suppressMessages(library(patchwork))
suppressMessages(library(ggpubr))
suppressMessages(library(plyr))
suppressMessages(library(dplyr))
suppressMessages(library(stringr))
suppressMessages(library(cowplot))
suppressMessages(library(ggplot2))
suppressMessages(library(RColorBrewer))
suppressMessages(library(scales))
suppressMessages(library(gProfileR))
suppressMessages(library(biomaRt))
suppressMessages(library(getopt))

```

## -----------------------------------------------------------------------------
## Load function
## -----------------------------------------------------------------------------

```{r}
source("/shared/ifbstor1/projects/scrna_gmp_rara/stage_babacar/scranseq/analysis/R_src/funForSeuratAnalysis.R")
```

## -----------------------------------------------------------------------------
## Command line args
## -----------------------------------------------------------------------------

```{r}
spec = matrix(c(
  'help',        'h', 0, "logical",   "Help about the program",
  "inputSeurat",  "i",1, "character", "input seurat object with cluster column names numclust",
  'outdir',     'o',1, "character", 'Outdir path (default ./)',
  "resolution", "r", 1, "character", "clustering resolution to use",
  #"SigList", "s", 1, "character", "Path to the directory with all the signatures (.txt)",
  #"clustName", "c", 1, "character", "List of cluster name sep by +",
  "logfc_threshold", "f", 1, "numeric", "Set a logFC threshold",
  "pval", "p", 1, "numeric", "Set a pval adj threshold",
  "assay", "a", 1, "character", "Chose default assay"
), byrow=TRUE, ncol=5);

opt = getopt(spec)
```

## Paths and options

```{r}
inputSeurat <- "/shared/projects/scrna_gmp_rara/stage_babacar/scranseq/analysis/output/RNA/Old_New_integration/rna_new.rds"
outdir <- "/shared/projects/scrna_gmp_rara/stage_babacar/scranseq/analysis/output/RNA/Old_New_integration/Analysis/"
assay <- "RNA"
#clustName <- "Neu2+NeuRA1+Rep+NeuRA2+Neu1+Neu3"



# Read seurat object
seurat <- readRDS(inputSeurat)

outdir <- outdir

# Set logFC threshold
if (is.null(opt$logfc_threshold)) {
  opt$logfc_threshold <- 0.25
}

# Set logFC threshold
if (is.null(opt$pval)) {
  opt$pval <- 0.05
}

# Define ref clustering
if(is.null(opt$resolution)){
  opt$resolution <- "integrated_snn_res.0.3"
}
#targetRes <- opt$resolution

# Define default assay
if(is.null(opt$assay)){
  opt$assay <- "RNA"
}

DefAssay <- assay
```

## ---------------------------------------------------------------------
## Color scheme
## ---------------------------------------------------------------------

```{r}
colorTreatment <- c("#664CFF", "#FF8000")
colorCluster <- c("#E69F00", "#CC79A7", "#0072B2", "#009E73", "#D55E00", "#56B4E9")
colorPhases <- c(brewer.pal(9,"RdPu"))[c(3,6,9)]
colorBranches <- c("#332288", "#117733", "#DDCC77", "#AA4499", "#88CCEE")
```

##-------------------------------------------------#
##       Data preparation : Clust name etc
##-------------------------------------------------#

```{r}
DefaultAssay(seurat) <- DefAssay

seurat@meta.data$phases <- factor(seurat@meta.data$phases, levels = c("G1_G0", "S", "G2_M"))

Idents(seurat) <- "predicted.id"
markers_cluster <- FindAllMarkers(seurat, logfc.threshold = opt$logfc_threshold)
markers_cluster_sig <- markers_cluster[markers_cluster$p_val_adj < opt$pval,]

# Rep markers
markers_Rep_sig = markers_cluster_sig[markers_cluster$cluster == "Rep", ]$gene
#cat("Rep markers : ", markers_Rep_sig)

#write.table(x = markers_cluster_sig, file = paste0(outdir, "Markers_table.tsv"), quote = F, col.names = T, row.names = F, sep = "\t")

```

##-----------------------------------------------------#
##   Looking for geneSet enrichment on cluster markers
##-----------------------------------------------------#

```{r}
enrich_clust <- list()
for(cluster in unique(markers_cluster$cluster)){
  print(cluster)
  #enrich_clust[[cluster]] <- gprofileR2::gost(markers_cluster[markers_cluster$cluster %in% cluster, "gene"],
                                                  #organism = "mmusculus",
                                                  #custom_bg = rownames(seurat),
                                                  #ordered_query = "none")
  
  markers_cluster_gost = markers_cluster[markers_cluster$cluster %in% cluster, "gene"]
  enrich_clust[[cluster]] <- gProfileR::gprofiler(query = markers_cluster_gost,
                                                  organism = "mmusculus",
                                                  custom_bg = rownames(seurat),
                                                  ordered_query = "none")
  enrich_clust[[cluster]]$cluster <- cluster
}
enrich_clust.df <- do.call("rbind", enrich_clust)

cat("top geneSet enrichment on cluster markers : ") 
head(enrich_clust.df)

#write.table(enrich_clust.df, file = paste0(outdir, "enrichmentTable.tsv"), sep = "\t", row.names = F, col.names = T, quote = F)

```

#-------------------------------------------------#
#              Basic UMAP plotting
#-------------------------------------------------#
dir.create(paste0(outdir, "UMAP/"), recursive = T)

```{r}
UMAP_Clust <- DimPlot(seurat, pt.size = 0.4, cols = colorCluster, group.by = "predicted.id") + theme(axis.title = element_blank(), axis.text = element_blank())
UMAP_treatment <- DimPlot(seurat, split.by = "sampleName", group.by = "sampleName", pt.size = 0.4, cols = colorTreatment) + theme(axis.title = element_blank(), axis.text = element_blank())

# UMAP treatment split
Idents(seurat) <- "sampleName"
seurat_control <- subset(seurat, idents = "RNA_Ctrl")
seurat_treated <- subset(seurat, idents = "RNA_RA")
Idents(seurat) <- "predicted.id"

UMAP_control <- DimPlot(seurat_control, split.by = "sampleName", group.by = "sampleName", pt.size = 0.4, cols = colorTreatment[1]) + theme(axis.title = element_blank(), axis.text = element_blank(), legend.position = "none")
UMAP_treated <- DimPlot(seurat_treated, split.by = "sampleName", group.by = "sampleName", pt.size = 0.4, cols = colorTreatment[2]) + theme(axis.title = element_blank(), axis.text = element_blank(), legend.position = "none")

UMAP_CellCycle <- DimPlot(seurat, pt.size = 0.4, cols = colorPhases, group.by = "phases") + theme(axis.title = element_blank(), axis.text = element_blank())
UMAP_CellCycle_split <- DimPlot(seurat, split.by = "sampleName", group.by = "phases", pt.size = 0.4, cols = colorPhases) + theme(axis.title = element_blank(), axis.text = element_blank())

#ggsave(paste0(outdir,"UMAP/", 'UMAP_CellCycle', '.png'), plot = UMAP_CellCycle, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 15, units = 'cm', dpi = 300)

#ggsave(paste0(outdir,"UMAP/", 'UMAP_split_CellCycle', '.png'), plot = UMAP_CellCycle_split, device = 'png', path = NULL,
       #scale = 1, width = 20, height = 15, units = 'cm', dpi = 300)
#ggsave(paste0(outdir,"UMAP/", 'UMAP_cluster', '.png'), plot = UMAP_Clust, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 15, units = 'cm', dpi = 300)

#ggsave(paste0(outdir, "UMAP/", 'UMAP_split_treatment', '.png'), plot = UMAP_treatment, device = 'png', path = NULL,
       #scale = 1, width = 20, height = 15, units = 'cm', dpi = 300)
#ggsave(paste0(outdir, "UMAP/", 'UMAP_treatment_Control', '.png'), plot = UMAP_control, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 15, units = 'cm', dpi = 300)
#ggsave(paste0(outdir, "UMAP/", 'UMAP_treatment_Treated', '.png'), plot = UMAP_treated, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 15, units = 'cm', dpi = 300)

UMAP_CellCycle
UMAP_Clust
UMAP_treatment
UMAP_control
UMAP_treated
```

#-------------------------------------------------#
# Generate the barplot for cell distrib in cluster
#-------------------------------------------------#

```{r}
dir.create(paste0(outdir, "BarPlot/"), recursive = T)
#BP of cell cycle distrib per cluster
summary_phases <- ddply(seurat@meta.data,~predicted.id + phases + sampleName, nrow)

bp_phasesClust_control <- ggplot(data.frame(summary_phases[summary_phases$sampleName %in% "RNA_Ctrl",]), aes(fill = phases,y = V1, x=predicted.id)) +
  geom_bar( stat="identity", position="fill")+
  scale_fill_manual( values= colorPhases)+
  scale_y_continuous(name = "Sample (%)", labels = c(0,25,50,75,100))+
  ylab(label = "")+xlab(label = "") + coord_flip()+
  theme(legend.title=element_blank()) + ggtitle("Control")

bp_phasesClust_treated <- ggplot(data.frame(summary_phases[summary_phases$sampleName %in% "RNA_RA",]), aes(fill = phases,y = V1, x=predicted.id)) +
  geom_bar( stat="identity", position="fill")+
  scale_fill_manual( values= colorPhases)+
  scale_y_continuous(name = "Sample (%)", labels = c(0,25,50,75,100))+
  ylab(label = "")+xlab(label = "") + coord_flip()+
  theme(legend.title=element_blank()) + ggtitle("Treated")

cellCycle_legend <- get_legend(bp_phasesClust_treated)

bp_grid_cellCycle <- plot_grid(bp_phasesClust_control + theme(legend.position = "none"), bp_phasesClust_treated + theme(legend.position = "none"))
bp_cellCycle_leg_grid <- plot_grid(bp_grid_cellCycle, cellCycle_legend,  rel_widths = c(3,0.4))

bp_cellCycle_leg_grid
#ggsave(paste0(outdir, "BarPlot/", 'BP_cellCycle_split', '.png'), plot = bp_cellCycle_leg_grid, device = 'png', path = NULL,
       #scale = 1, width = 20, height = 20, units = 'cm', dpi = 300)

CellPerClust <- ddply(seurat@meta.data,~predicted.id + sampleName, nrow)

bp_rawCellCount_control <- ggplot(data.frame(CellPerClust[CellPerClust$sampleName %in% "RNA_Ctrl",]), aes(y = V1, x=predicted.id)) +
  geom_bar( stat="identity") + coord_flip() + ylab("nCells") + expand_limits(y = c(NULL, 3000))
bp_rawCellCount_treated <- ggplot(data.frame(CellPerClust[CellPerClust$sampleName %in% "RNA_RA",]), aes(y = V1, x=predicted.id)) +
  geom_bar( stat="identity") + coord_flip() + ylab("nCells") + expand_limits(y = c(NULL, 3000))

bp_rawCellCount_control
bp_rawCellCount_treated

#ggsave(paste0(outdir, "BarPlot/", 'BP_raw_cellCount_control', '.png'), plot = bp_rawCellCount_control, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 20, units = 'cm', dpi = 300)
#ggsave(paste0(outdir, "BarPlot/", 'BP_raw_cellCount_treated', '.png'), plot = bp_rawCellCount_treated, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 20, units = 'cm', dpi = 300)

bp_control_raw_cycle <- plot_grid(bp_phasesClust_control + theme(legend.position = "none"), bp_rawCellCount_control + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + ggtitle(" "), rel_widths = c(2,1))
bp_treated_raw_cycle <- plot_grid(bp_phasesClust_treated + theme(legend.position = "none"), bp_rawCellCount_treated + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + ggtitle(" "), rel_widths = c(2,1))

bp_all_grid_noleg <- plot_grid(bp_control_raw_cycle, bp_treated_raw_cycle, nrow = 2)
bp_all_grid_leg <- plot_grid(bp_all_grid_noleg, cellCycle_legend, rel_widths = c(3,0.4))

bp_treated_raw_cycle
#ggsave(paste0(outdir, "BarPlot/", 'BP_grid_cellCycle_cellCount', '.png'), plot = bp_treated_raw_cycle, device = 'png', path = NULL,
       #scale = 1, width = 20, height = 20, units = 'cm', dpi = 300)

# BP proportion treatment per cluster
clustersampleName <- ddply(seurat@meta.data,~predicted.id + sampleName,nrow)
propExpect <- table(seurat@meta.data$sampleName)/length(seurat@meta.data$sampleName)[]


enrich_clust <- getEnrichPopClust(hspc.combined = seurat, Xname = "RNA_Ctrl", Yname = "RNA_RA", colorX = colorTreatment[1], colorY = colorTreatment[2], metaCol = "sampleName", clustCol = "predicted.id")


clustersampleName$sampleName <- factor(clustersampleName$sampleName, levels = c("RNA_RA", "RNA_Ctrl"))
clustersampleName$condition <- factor(clustersampleName$sampleName, levels = c("RNA_Ctrl", "RNA_RA"))

bp_treatmentClust1 <- ggplot(data.frame(clustersampleName), aes(fill = sampleName,y = V1, x=predicted.id)) +
  geom_bar( stat="identity", position="fill")+
  scale_fill_manual( values= rev(colorTreatment))+
  scale_y_continuous(name = "Sample (%)", labels = c(0,25,50,75,100))+
  ylab(label = "")+xlab(label = "") + theme_classic() + coord_flip()+
  theme(legend.title=element_blank(), axis.text.y = element_text(colour = enrich_clust[,"color"])) +
  geom_hline(yintercept = propExpect[1])

bp_treatmentClust1
#ggsave(paste0(outdir, "BarPlot/",'BP_treatment_fill', '.png'), plot = bp_treatmentClust1, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 20, units = 'cm', dpi = 300)


bp_treatmentClust2 <- ggplot(data.frame(clustersampleName), aes(fill = condition,y = V1, x=predicted.id)) +
  #geom_bar( stat="identity", position="fill")+
  geom_bar( stat="identity", position = position_dodge2())+
  scale_fill_manual( values= colorTreatment)+
  ylab(label = "nCells")+xlab(label = "") +
  theme_classic() + coord_flip()
  #theme(legend.title=element_blank(), axis.text.y = element_text(colour = enrich_clust[,"color"])) 
  #geom_hline(yintercept = propExpect[1])

bp_treatmentClust2
#ggsave(paste0(outdir, "BarPlot/",'BP_treatment_position_dodge', '.png'), plot = bp_treatmentClust2, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 20, units = 'cm', dpi = 300)

# BP cell number distribution per cluster 
CellPerClust <- ddply(seurat@meta.data,~predicted.id, nrow)

bp_rawTreatmentClust <- ggplot(data.frame(CellPerClust), aes(y = V1, x=predicted.id)) +
  geom_bar( stat="identity") + theme_classic() + coord_flip() + ylab("nCells") + xlab("")

bp_rawTreatmentClust
#ggsave(paste0(outdir, "BarPlot/",'BP_raw_treatment', '.png'), plot = bp_rawTreatmentClust, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 20, units = 'cm', dpi = 300)

# Arranging BP grid
treatment_legend <- get_legend(bp_treatmentClust2)

bp_grid1 <- plot_grid(bp_treatmentClust1 + theme(legend.position = "none", axis.line.y = element_blank(), 
                     axis.ticks.y = element_blank()), 
                     bp_treatmentClust2  + theme(legend.position = "none", axis.text.y = element_blank(), 
                     axis.title.y = element_blank(), axis.line.y = element_blank(), axis.ticks.y = element_blank()),
                     treatment_legend, rel_widths = c(2,2,0.5), nrow = 1)

bp_grid1
#ggsave(paste0(outdir, "BarPlot/",'BP_grid1_treatment_dodge', '.png'), plot = bp_grid1, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 20, units = 'cm', dpi = 300)


bp_grid2 <- plot_grid(bp_treatmentClust1 + theme(legend.position = "none", axis.line.y = element_blank(), 
                     axis.ticks.y = element_blank()), 
                     bp_rawTreatmentClust + theme(axis.text.y = element_blank(), axis.title.y = element_blank(), 
                     axis.line.y = element_blank(), 
                     axis.ticks.y = element_blank()), 
                     treatment_legend, rel_widths = c(2,1,0.5), nrow = 1)

bp_grid2
#ggsave(paste0(outdir, "BarPlot/",'BP_grid2_treatment_dodge', '.png'), plot = bp_grid2, device = 'png', path = NULL,
       #scale = 1, width = 15, height = 20, units = 'cm', dpi = 300)
```

## results saving
```{r}
Idents(seurat) <- "predicted.id"
DefaultAssay(seurat) <- DefAssay
#saveRDS(seurat, paste0(outdir, "seurat_annotated", ".rds"))
```


## -----------------------------------------------------------------------------
############# hcG-PLZF-RARA STATICAL ANALYSIS#############
## -----------------------------------------------------------------------------

B- hcG-PLZF-RARA STATICAL ANALYSIS


```{r}
## Directoies

inputSeurat <- "/shared/projects/scrna_gmp_rara/stage_babacar/scranseq/analysis/output/RNA/Old_New_integration/rna_new.rds"

outdir <- "/shared/projects/scrna_gmp_rara/stage_babacar/scranseq/analysis/output/RNA/Old_New_integration/Analysis/hcG-PLZF-RARA_stats"

# Assay choice
assay <- "RNA"
seurat <- readRDS(inputSeurat)
rna_new = seurat

outdir <- outdir

# colors as factor
colorTreatment <- c("#664CFF", "#FF8000")
colorCluster <- c("#E69F00", "#CC79A7", "#0072B2", "#009E73", "#D55E00", "#56B4E9")
colorPhases <- c(brewer.pal(9,"RdPu"))[c(3,6,9)]
colorBranches <- c("#332288", "#117733", "#DDCC77", "#AA4499", "#88CCEE")

DefaultAssay(seurat) <- DefAssay


# compare mean expression of hcG-PLZF-RARA per cluster per treatment

p1 = VlnPlot(rna_new, feature = "PLZF-RARA-hcG", pt.size = 0, group.by = "predicted.id", split.by = "sampleName", 
             cols = rep(colorTreatment)) 
# Add pairwise comparisons p-value
p1 = p1 + stat_compare_means(method = "anova", label = "p.signif") 

#png(paste(outdir,"/hcG-PLZF-RARA_MeanComparisonPerClusterPerTreatment.png", sep = "") ,res = 300,height=6,width =10,units = "in") # 
grid.arrange(p1)
#dev.off()

# compare mean expression of hcG-PLZF-RARA per treatment

#my_comparisons <- list(c("RNA_Ctrl", "RNA_RA"))
p2 = VlnPlot(rna_new, feature = "PLZF-RARA-hcG", pt.size = 0, group.by = "sampleName", cols = rep(colorTreatment)) + NoLegend()
p2 = p2 + stat_compare_means(method = "t.test", label = "p.signif")

#png(paste(outdir,"/hcG-PLZF-RARA_MeanComparisonPerTreatment.png", sep = "") ,res = 300,height=6,width =10,units = "in")
grid.arrange(p2)
#dev.off()


# compare mean expression of hcG-PLZF-RARA per cluster

#my_comparisons <- list(c("Rep", "Neu1", "Neu2", "Neu3", "NeuRA1", "NeuRA2"))
p3 = VlnPlot(rna_new, feature = "PLZF-RARA-hcG", pt.size = 0, group.by = "predicted.id", cols = rep(colorCluster)) 
p3 = p3 + stat_compare_means(method = "anova", label = "p.signif")

#png(paste(outdir,"/hcG-PLZF-RARA_MeanComparisonPerCluster.png", sep = "") ,res = 300,height=6,width =10,units = "in") # 
grid.arrange(p3)
#dev.off()

# save plots
#png(paste(outdir,"/hcG-PLZF-RARA_MeanComparison.png", sep = "") ,res = 300,height=6,width =10,units = "in") # 
#grid.arrange(p1, p2, p3, nrow = 1)
#dev.off()

```

